---
title: "Models"
output: html_document
date: "2024-05-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lattice)
library(gstat)
library(sp)
library(ggplot2)
library(scales)
library(plotly)
library(raster)
library(dismo)
library(sf)
library(tidyverse)
library(hrbrthemes)
library(terra)
library(RColorBrewer)
#netCDF packages
library(ncdf4)
library(CFtime)
library(stars)
library(here)
```

##Locations 
```{r}
data_loc <- here("Data")
pca_loc <- here("Data", "PCA")
outputs_loc <- here("Outputs")
averages_loc <- here("Outputs", "AvgRasters")
```

##BNF first 
```{r}
loc <- here("Data", "TRENDYN")
CP_BNF <- rast(here(loc, "CABLE-POP/fBNF_annually.nc"))
CLASSIC_BNF <- rast(here(loc,"CLASSIC/fBNF_annually.nc"))
CLM5_BNF <- rast(here(loc,"CLM5.0/fBNF_annually.nc"))
DLEM_BNF <- rast(here(loc,"DLEM/fBNF_annually.nc"))
ISAM_BNF <- rast(here(loc,"ISAM/fBNF_annually.nc"))
JSBACH_BNF <- rast(here(loc,"JSBACH/fBNF_annually.nc"))
JULES_ES_BNF <- rast(here(loc,"JULES-ES/fBNF_annually.nc"))
LPJ_GUESS_BNF <- rast(here(loc,"LPJ-GUESS/fBNF_annually.nc"))
LPX_Bern_BNF <- rast(here(loc,"LPX-Bern/fBNF_annually.nc"))
OCNv2_BNF <- rast(here(loc,"OCNv2/fBNF_annually.nc"))
ORCHIDEEv3_BNF <- rast(here(loc,"ORCHIDEEv3/fBNF_annually.nc"))
```

```{r}
head(names(CLASSIC_BNF))
```

##as a raster
```{r}
class(CLASSIC_BNF)
```

```{r}
plot(CLASSIC_BNF)
```
##change coordinates to normal
```{r}
CLASSIC_BNF <- rotate(CLASSIC_BNF)
CLASSIC_BNF
```
```{r}
plot(CLASSIC_BNF)
```











## read as a netcdf file
```{r}
Classic_Nloss_nc <- ncdf4::nc_open(here(loc, "ORCHIDEEv3/fNloss_annually.nc"))
Classic_Nloss_nc
```

```{r}
#print(Classic_BNF_nc)
```




##Average over the selected timestamps 
```{r}
data_var <- ncvar_get(Classic_Nloss_nc, "fNloss")
time_var <- ncvar_get(Classic_Nloss_nc, "time")

lat <- ncvar_get(Classic_Nloss_nc, "latitude")
lon <- ncvar_get(Classic_Nloss_nc, "longitude")
```



```{r}
nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)
```


```{r}
selected_time_steps <- c(149:171)
```

```{r}
rasters <- list()
```



```{r}
# Extract data for the selected time steps and create raster layers
for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}
```

```{r}
# Combine the rasters into a raster brick
raster_brick <- brick(rasters)
#print(raster_brick)
```

```{r}
# Calculate the average
average_raster <- calc(raster_brick, mean)
```

```{r}
average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

# Define the output file path
output_file <- here(averages_loc,"ORCHIDEEv3_Nloss_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)
```





CP_BNF <- rast(here(loc, "CABLE-POP/fBNF_annually.nc"))
CLASSIC_BNF <- rast(here(loc,"CLASSIC/fNloss_annually.nc"))
CLM5_BNF <- rast(here(loc,"CLM5.0/fNloss_annually.nc"))
DLEM_BNF <- rast(here(loc,"DLEM/fNloss_annually.nc"))
ISAM_BNF <- rast(here(loc,"ISAM/fNloss_annually.nc"))
JSBACH_BNF <- rast(here(loc,"JSBACH/fNloss_annually.nc"))
JULES_ES_BNF <- rast(here(loc,"JULES-ES/fNloss_annually.nc"))
LPJ_GUESS_BNF <- rast(here(loc,"LPJ-GUESS/fNloss_annually.nc"))
LPX_Bern_BNF <- rast(here(loc,"LPX-Bern/fNloss_annually.nc"))
OCNv2_BNF <- rast(here(loc,"OCNv2/fNloss_annually.nc"))

ORCHIDEEv3_BNF <- rast(here(loc,"ORCHIDEEv3/fNloss_annually.nc"))


```{r}
Classic_bnf_nc <- ncdf4::nc_open(here(loc, "CLASSIC/fBNF_annually.nc"))

data_var <- ncvar_get(Classic_bnf_nc, "fBNF")
time_var <- ncvar_get(Classic_bnf_nc, "time")

lat <- ncvar_get(Classic_bnf_nc, "latitude")
lon <- ncvar_get(Classic_bnf_nc, "longitude")

nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)

selected_time_steps <- c(150:170)

rasters <- list()

for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}

raster_brick <- brick(rasters)

average_raster <- calc(raster_brick, mean)




average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

values(average_raster_units)

# Define the output file path
output_file <- here(averages_loc,"Classic_bnf_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)

```

##N loss


```{r}
#averages_loc_nloss <- here("Outputs", "AvgRasters", "Nloss")

Classic_nloss_nc <- ncdf4::nc_open(here(loc, "ORCHIDEEv3/fNloss_annually.nc"))

data_var <- ncvar_get(Classic_nloss_nc, "fNloss")
time_var <- ncvar_get(Classic_nloss_nc, "time")

lat <- ncvar_get(Classic_nloss_nc, "latitude")
lon <- ncvar_get(Classic_nloss_nc, "longitude")

nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)

selected_time_steps <- c(150:170)

rasters <- list()

for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}

raster_brick <- brick(rasters)

average_raster <- calc(raster_brick, mean)




average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

values(average_raster_units)

# Define the output file path
output_file <- here(averages_loc_nloss,"ORCHIDEEv3_Nloss_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)

```




























##All means and Variance Denit
NOT WORKING
```{r}
# Define the directory containing your raster files
rasters_bnf <- here("Outputs", "AvgRasters", "BNF")

# List all .tif files in the directory
raster_files <- list.files(rasters_bnf, pattern = "\\.tif$", full.names = TRUE)

# Create a SpatRaster stack from the list of files
raster_stack <- rast(raster_files)

print(raster_stack)
plot(raster_stack)

#values(raster_stack)

#writeRaster(raster_stack, file = "Stack_TRENDYN_BNF.tif", format = "GTiff", overwrite = TRUE)

# Calculate the variance of the raster stack
#variance <- calc(raster_stack, var)

# Plot the variance raster
#plot(variance)

# Write the adjusted average raster to a GeoTIFF file
#writeRaster(variance, file = "Variance_TRENDYN_Nloss.tif", format = "GTiff", overwrite = TRUE)
```

```{r}

rasters_bnf <- here("Outputs", "AvgRasters", "BNF")
raster_files <- c(
  here(rasters_bnf, "Classic_bnf_Avg.tif"),
  here(rasters_bnf, "Clm5_bnf_Avg.tif"),
  here(rasters_bnf, "CP_bnf_Avg.tif"),
  here(rasters_bnf, "Dlem_bnf_Avg.tif"),
  here(rasters_bnf, "Isam_bnf_Avg.tif"),
  here(rasters_bnf, "Jsbach_bnf_Avg.tif"),
  here(rasters_bnf, "Jules_es_bnf_Avg.tif"),
  here(rasters_bnf, "Lpj_guess_bnf_Avg.tif"),
  here(rasters_bnf, "Lpx_bern_bnf_Avg.tif"),
  here(rasters_bnf, "Ocnv2_bnf_Avg.tif"),
  here(rasters_bnf, "Orchideev3_bnf_Avg.tif")
)


raster_stack <- rast(raster_files)

print(raster_stack)
plot(raster_stack)

stack_variance <- app(raster_stack, fun = function(x) var(x, na.rm = TRUE))

print(stack_variance)
plot(stack_variance)

# Define the output file path
output_file <- here(averages_loc,"BNF_Variance.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(stack_variance, output_file, overwrite = TRUE)



mean_raster <- app(raster_stack, fun = function(x) mean(x, na.rm = TRUE))
print(mean_raster)
plot(mean_raster)
# Define the output file path
output_file <- here(averages_loc,"BNF_mean.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(mean_raster, output_file, overwrite = TRUE)

sd_raster <- app(raster_stack, fun = function(x) sd(x, na.rm = TRUE))
print(sd_raster)
plot(sd_raster)
output_file <- here(averages_loc,"BNF_SD.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(sd_raster, output_file, overwrite = TRUE)

cv_raster <- sd_raster / mean_raster
print(cv_raster)
plot(cv_raster)
output_file <- here(averages_loc,"BNF_CV.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(cv_raster, output_file, overwrite = TRUE)
```


```{r}
#averages_loc_nloss <- here("Outputs", "AvgRasters", "Nloss")
raster_files <- c(
  here(averages_loc_nloss, "Classic_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Clm5_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Dlem_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Isam_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Jsbach_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Jules_es_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Lpj_guess_Nloss_Avg.tif"),
  here(averages_loc_nloss, "Lpx_bern_Nloss_Avg.tif"),
  here(averages_loc_nloss, "OCNv2_Nloss_Avg.tif"),
  here(averages_loc_nloss, "ORCHIDEEv3_Nloss_Avg.tif")
)


raster_stack <- rast(raster_files)

print(raster_stack)
plot(raster_stack)

stack_variance <- app(raster_stack, fun = function(x) var(x, na.rm = TRUE))

print(stack_variance)
plot(stack_variance)

# Define the output file path
output_file <- here(averages_loc,"Nloss_Variance.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(stack_variance, output_file, overwrite = TRUE)



mean_raster <- app(raster_stack, fun = function(x) mean(x, na.rm = TRUE))
print(mean_raster)
plot(mean_raster)
# Define the output file path
output_file <- here(averages_loc,"Nloss_mean.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(mean_raster, output_file, overwrite = TRUE)

sd_raster <- app(raster_stack, fun = function(x) sd(x, na.rm = TRUE))
print(sd_raster)
plot(sd_raster)
output_file <- here(averages_loc,"Nloss_SD.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(sd_raster, output_file, overwrite = TRUE)

cv_raster <- sd_raster / mean_raster
print(cv_raster)
plot(cv_raster)
output_file <- here(averages_loc,"Nloss_CV.tif")
# Write the adjusted average raster to a GeoTIFF file
writeRaster(cv_raster, output_file, overwrite = TRUE)
```







##########################################################################################################

##THIS IS NOT WORKING FOR MYSTERY REASONS
```{r}


# Convert the RasterStack to a SpatRaster
spat_raster_stack <- rast(raster_stack)
plot(spat_raster_stack)

# Calculate the variance across the SpatRaster stack
variance_raster <- app(spat_raster_stack, fun = function(x) var(x, na.rm = TRUE))

# Plot the variance raster
plot(variance_raster)

#output_file <- "Variance_TRENDYN_Nloss.tif"
#writeRaster(variance_raster, output_file, overwrite = TRUE)
```

# Denit - mean, stdev, cv
```{r}
# Calculate the mean across the raster stack, removing NA values
mean_raster <- app(spat_raster_stack, fun = function(x) mean(x, na.rm = TRUE))
#output_file <- "Mean_TRENDYN_Nloss.tif"
#writeRaster(mean_raster, output_file, overwrite = TRUE)

# Calculate the standard deviation across the raster stack, removing NA values
sd_raster <- app(spat_raster_stack, fun = function(x) sd(x, na.rm = TRUE))
#output_file <- "SD_TRENDYN_Nloss.tif"
#writeRaster(sd_raster, output_file, overwrite = TRUE)

# Calculate the coefficient of variation (CV)
cv_raster <- sd_raster / mean_raster
#output_file <- "CV_TRENDYN_Nloss.tif"
#writeRaster(cv_raster, output_file, overwrite = TRUE)

# Plot the CV raster
plot(sd_raster)
```









##Variance BNF
```{r}
# Define the directory containing your raster files
rasters_nloss <- here("Outputs", "AvgRasters", "Nloss")

# List all .tif files in the directory
raster_files <- list.files(rasters_nloss, pattern = "\\.tif$", full.names = TRUE)

# Create a SpatRaster stack from the list of files
raster_stack <- rast(raster_files)

mean_raster <- app(raster_stack, fun = function(x) mean(x, na.rm = TRUE))

plot(mean_raster)
# Calculate the variance of the raster stack
#variance <- calc(raster_stack, var)

# Plot the variance raster
#plot(variance)

# Write the adjusted average raster to a GeoTIFF file
#writeRaster(variance, file = "Variance_TRENDYN_BNF.tif", format = "GTiff", overwrite = TRUE)
```

```{r}
# Convert the RasterStack to a SpatRaster
spat_raster_stack <- rast(raster_stack)

# Calculate the variance across the SpatRaster stack
variance_raster <- app(spat_raster_stack, fun = function(x) var(x, na.rm = TRUE))

# Plot the variance raster
plot(variance_raster)

#output_file <- "Variance_TRENDYN_BNF.tif"
#writeRaster(variance_raster, output_file, overwrite = TRUE)
```

# Denit - mean, stdev, cv
```{r}
# Calculate the mean across the raster stack, removing NA values
mean_raster <- app(spat_raster_stack, fun = function(x) mean(x, na.rm = TRUE))
output_file <- here("Outputs", "Mean_TRENDYN_BNF.tif")
writeRaster(mean_raster, output_file, overwrite = TRUE)

# Calculate the standard deviation across the raster stack, removing NA values
sd_raster <- app(spat_raster_stack, fun = function(x) sd(x, na.rm = TRUE))
output_file <- here("Outputs","SD_TRENDYN_BNF.tif")
writeRaster(sd_raster, output_file, overwrite = TRUE)

# Calculate the coefficient of variation (CV)
cv_raster <- sd_raster / mean_raster
output_file <- here("Outputs","CV_TRENDYN_BNF.tif")
writeRaster(cv_raster, output_file, overwrite = TRUE)

# Plot the CV raster
plot(sd_raster)
```

##########################################################################################################








##PLotting density curves with TRENDY data

```{r}
bnf <- read.csv(here(data_loc,'BNF_TRENDYN.csv'), header= TRUE)
bnf_no_NA <- bnf %>% drop_na(Variance)

denit <- read.csv(here(data_loc,'Nloss_TRENDYN.csv'), header= TRUE)
denit_no_NA <- denit %>% drop_na(Variance)
```

##BNF - Variance 

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
var_bnf_global <- brick(here(averages_loc, "BNF_Variance.tif"))
var_bnf_global_df <- as.data.frame(rasterToPoints(var_bnf_global,spatial = TRUE))
```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
var_bnf <- brick(here(outputs_loc, "Tropics_BNF_Variance.tif"))
var_bnf_df <- as.data.frame(rasterToPoints(var_bnf,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,40)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = var_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = Variance, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model variance - no Sahara", x = "g N M-2 yr -1", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
x_axis_extent <- c(0,40)

# Specify the colors for each dataset
color_palette <- c("BNF - Global" = "grey", "BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = var_bnf_global_df, aes(x = lyr.1, fill = "BNF - Global"), alpha = 0.5) +
  geom_density(data = var_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = Variance, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model variance", x = "g N M-2 yr -1", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```





#BNF Density curves - CV

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_bnf_global <- brick(here(averages_loc, "BNF_CV.tif"))
cv_bnf_global_df <- as.data.frame(rasterToPoints(cv_bnf_global,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,5)

# Specify the colors for each dataset
color_palette <- c("BNF - Global" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_bnf_global_df, aes(x = lyr.1, fill = "BNF - Global"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = CV, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model CV", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

#BNF Density curves - CV - No Sahara

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_bnf <- brick(here(outputs_loc,"Tropics_BNF_CV.tif"))
cv_bnf_df <- as.data.frame(rasterToPoints(cv_bnf,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = CV, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model CV - No Sahara", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("BNF - Global" = "grey", "BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_bnf_global_df, aes(x = lyr.1, fill = "BNF - Global"), alpha = 0.5) +
  geom_density(data = cv_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = CV, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of modeled BNF CV", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.85))  # Adjust these values for the desired position

```






SD_TRENDYN_BNF.tif
```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
sd_bnf <- brick(here(outputs_loc,"Tropics_BNF_SD.tif"))
sd_bnf_df <- as.data.frame(rasterToPoints(sd_bnf,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = sd_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = SD, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model StDev - No Sahara", x = "Standard Deviation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
sd_bnf_global <- brick(here(averages_loc,"BNF_SD.tif"))
sd_bnf_global_df <- as.data.frame(rasterToPoints(sd_bnf_global,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Global" = "grey","BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = sd_bnf_global_df, aes(x = lyr.1, fill = "BNF - Global"), alpha = 0.5) +
  geom_density(data = sd_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = SD, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model StDev", x = "Standard Deviation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```




##BNF TRENDY Mean
```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
mean_bnf <- brick(here(outputs_loc,"Tropics_BNF_Mean.tif"))
mean_bnf_df <- as.data.frame(rasterToPoints(mean_bnf,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = mean_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = Mean, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model Mean - No Sahara", x = "Mean all models", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
mean_bnf_global <- brick(here(outputs_loc,"Mean_TRENDYN_BNF.tif"))
mean_bnf_global_df <- as.data.frame(rasterToPoints(mean_bnf_global,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Global" = "grey","BNF - Tropics" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = mean_bnf_global_df, aes(x = lyr.1, fill = "BNF - Global"), alpha = 0.5) +
  geom_density(data = mean_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = Mean, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model mean", x = "Mean of all models", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```










##N loss TRENDY

##N loss - Variance


```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
Var_denit_global <- brick(here(averages_loc,"Nloss_Variance.tif"))
Var_denit_global_df <- as.data.frame(rasterToPoints(cv_denit_global,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,50)

# Specify the colors for each dataset
color_palette <- c("N loss - Global" = "grey", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = Var_denit_global_df, aes(x = lyr.1, fill = "N loss - Global"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = Var, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model variance - no Sahara", x = "g N m-2 yr -1", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
Var_denit <- brick(here(outputs_loc,"Tropics_Nloss_Variance.tif"))
Var_denit_df <- as.data.frame(rasterToPoints(Var_denit,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,50)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = Var_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = Variance, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model variance - no Sahara", x = "g N m-2 yr -1", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
x_axis_extent <- c(0,50)

# Specify the colors for each dataset
color_palette <- c("N loss - Global" = "grey", "N loss - Tropics" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = Var_denit_global_df, aes(x = lyr.1, fill = "N loss - Global"), alpha = 0.5) +
  geom_density(data = Var_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = Var, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model variance", x = "g N m-2 yr -1", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

#Denit Density curves - CV

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_denit_global <- brick(here(averages_loc,"Nloss_CV.tif"))
cv_denit_global_df <- as.data.frame(rasterToPoints(cv_denit_global,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "yellow", "Denit - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_denit_global_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = CV, fill = "Denit - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of Denit model CV", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


#Denit Density curves - CV - No Sahara

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_denit <- brick(here(outputs_loc,"Tropics_Nloss_CV.tif"))
cv_denit_df <- as.data.frame(rasterToPoints(cv_denit,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "yellow", "Denit - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = CV, fill = "Denit - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of Denit model CV - No Sahara", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("N loss - Global" = "grey", "N loss - Tropics" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_denit_global_df, aes(x = lyr.1, fill = "N loss - Global"), alpha = 0.5) +
  geom_density(data = cv_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = CV, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of modeled N loss CV", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.85))  # Adjust these values for the desired position

```





#Denit Density curves - StDev

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
sd_denit <- brick(here(outputs_loc,"Tropics_Nloss_SD.tif"))
sd_denit_df <- as.data.frame(rasterToPoints(sd_denit,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "yellow", "Denit - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = sd_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = SD, fill = "Denit - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of Denit model StDev - No Sahara", x = "Standard Deviation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
sd_denit_global <- brick(here(outputs_loc,"SD_TRENDYN_Nloss.tif"))
sd_denit_global_df <- as.data.frame(rasterToPoints(sd_denit_global,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("N loss - Global" = "grey", "N loss - Tropics" = "yellow", "Denit - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = sd_denit_global_df, aes(x = lyr.1, fill = "N loss - Global"), alpha = 0.5) +
  geom_density(data = sd_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = SD, fill = "Denit - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of Denit model StDev", x = "Standard Deviation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


##N loss TRENDY Mean

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
mean_nloss <- brick(here(outputs_loc,"Tropics_Nloss_Mean.tif"))
mean_nloss_df <- as.data.frame(rasterToPoints(mean_nloss,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = mean_nloss_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = Mean, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model Mean - No Sahara", x = "Mean all models", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
mean_nloss_global <- brick(here(outputs_loc,"Mean_TRENDYN_Nloss.tif"))
mean_nloss_global_df <- as.data.frame(rasterToPoints(mean_nloss_global,spatial = TRUE))
```

```{r}
x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("N loss - Global" = "grey", "N loss - Tropics" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = mean_nloss_global_df, aes(x = lyr.1, fill = "N loss - Global"), alpha = 0.5) +
  geom_density(data = mean_nloss_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = Mean, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model Mean", x = "Mean all models", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


##EBF only 
```{r}
bnf <- read.csv(here(data_loc,'BNF_TRENDYN.csv'), header= TRUE)
#bnf_no_NA <- bnf %>% drop_na(Var)
bnf_ebf <- bnf[bnf$Biome_rough == 'EBF', ]

denit <- read.csv(here(data_loc,'Nloss_TRENDYN.csv'), header= TRUE)
#denit_no_NA <- denit %>% drop_na(Var)
denit_ebf <- denit[denit$Biome_rough == 'EBF', ]
```





```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_bnf_ebf <- brick(here(outputs_loc, "EBF_BNF_CV.tif"))
cv_bnf_ebf_df <- as.data.frame(rasterToPoints(cv_bnf_ebf,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "grey","BNF - EBF" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = cv_bnf_ebf_df, aes(x = lyr.1, fill = "BNF - EBF"), alpha = 0.5) +
  geom_density(data = bnf_ebf, aes(x = CV, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model CV - Evergreen broadleaf forests", x = "Coefficient of variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_nloss_ebf <- brick(here(outputs_loc, "EBF_Nloss_CV.tif"))
cv_nloss_ebf_df <- as.data.frame(rasterToPoints(cv_nloss_ebf,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "grey","N loss - EBF" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = cv_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = cv_nloss_ebf_df, aes(x = lyr.1, fill = "N loss - EBF"), alpha = 0.5) +
  geom_density(data = denit_ebf, aes(x = CV, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model CV - Evergreen broadleaf forests", x = "Coefficient of variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
sd_bnf_ebf <- brick(here(outputs_loc, "EBF_BNF_SD.tif"))
sd_bnf_ebf_df <- as.data.frame(rasterToPoints(sd_bnf_ebf,spatial = TRUE))
```

```{r}

x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "grey","BNF - EBF" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = sd_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = sd_bnf_ebf_df, aes(x = lyr.1, fill = "BNF - EBF"), alpha = 0.5) +
  geom_density(data = bnf_ebf, aes(x = SD, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model SD - Evergreen broadleaf forests", x = "Standard deviation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
sd_nloss_ebf <- brick(here(outputs_loc, "EBF_Nloss_CV.tif"))
sd_nloss_ebf_df <- as.data.frame(rasterToPoints(sd_nloss_ebf,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,5)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "grey","N loss - EBF" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = sd_denit_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = sd_nloss_ebf_df, aes(x = lyr.1, fill = "N loss - EBF"), alpha = 0.5) +
  geom_density(data = denit_ebf, aes(x = SD, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model SD - Evergreen broadleaf forests", x = "Standard deviation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
mean_bnf_ebf <- brick(here(outputs_loc, "EBF_BNF_Mean.tif"))
mean_bnf_ebf_df <- as.data.frame(rasterToPoints(mean_bnf_ebf,spatial = TRUE))
```

```{r}

x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "grey", "BNF - EBF" = "yellow", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = mean_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = mean_bnf_ebf_df, aes(x = lyr.1, fill = "BNF - EBF"), alpha = 0.5) +
  geom_density(data = bnf_ebf, aes(x = Mean, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of BNF model Mean - Evergreen broadleaf forests", x = "Mean of all model BNF rates", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```


```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
mean_nloss_ebf <- brick(here(outputs_loc, "EBF_BNF_Mean.tif"))
mean_nloss_ebf_df <- as.data.frame(rasterToPoints(mean_nloss_ebf,spatial = TRUE))
```

```{r}

x_axis_extent <- c(0,7.5)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "grey", "N loss - EBF" = "yellow", "N loss - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
ggplot() +
  geom_density(data = mean_nloss_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = mean_nloss_ebf_df, aes(x = lyr.1, fill = "N loss - EBF"), alpha = 0.5) +
  geom_density(data = denit_ebf, aes(x = Mean, fill = "N loss - Sampled points"), alpha = 0.5) +
  labs(title = "Density Plot of N loss model Mean - Evergreen broadleaf forests", x = "Mean of all model N loss rates", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(legend.position = c(0.83, 0.86))  # Adjust these values for the desired position

```























##GAM
```{r}
library(mgcv)
```
```{r}
s <- scale(bnf$Elevation, center=TRUE, scale = TRUE)
```


```{r}
rand <- read.csv(here(data_loc,'RandPoints_CV_TRENDY.csv'), header= TRUE)

rand_bnf <- rand %>% dplyr::select(-c(OID.., Lat, Long, Type, PPT, MAT, AET, IGBP, Seasonality, GPP, NPP, NEE, Ndep2050, Slope, Elevation, Roughness, SoilOC, SoilN, SoilP, SoilpH, DtB1, CV_Nloss1))
rand_bnf <- rand_bnf %>% drop_na(CV_BNF1)

rand_Nloss <- rand %>% dplyr::select(-c(OID.., Lat, Long, Type, PPT, MAT, AET, IGBP, Seasonality, GPP, NPP, NEE, Ndep2050, Slope, Elevation, Roughness, SoilOC, SoilN, SoilP, SoilpH, DtB1, CV_BNF1))
rand_Nloss <- rand_Nloss %>% drop_na(CV_Nloss1)
```


##GAM - BNF
```{r}
tmp_sdm<-gam(PresAbs~ CV_BNF1,family = binomial, data=rand_bnf, select = TRUE, method="GCV.Cp")
#Then when you want to predict, you just:
Gam_pred <- predict.gam(tmp_sdm, newdata=data, type="response")
Gam_pred <- as.data.frame(Gam_pred)
#predictions for presence absence at each point based only on PPT and MAT
```

```{r}
hist(Gam_pred$Gam_pred)
```























