---
title: "TropicalBlindSpotsCode"
output: html_document
date: "2025-05-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lattice)
library(gstat)
library(stats)
library(sp)
library(ggplot2)
library(scales)
library(plotly)
library(raster)
library(dismo)
library(sf)
library(tidyverse)
library(hrbrthemes)
library(terra)
library(RColorBrewer)
#PCA packages
library(corrplot)
library(factoextra) 
library(FactoMineR)
library(vegan)
library(mgcv)
library(ggrepel)
library(caret)
#netCDF packages
library(ncdf4)
library(CFtime)
library(stars)
library(here)
```

##Locations 
```{r}
data_loc <- here("Data", "TBS_Datasets")
pca_loc <- here("Data", "PCA")
outputs_loc <- here("Outputs")
averages_loc <- here("Outputs", "AvgRasters")
rast_loc <- here("Data", "ClippedGloballyGridded")
figure_loc <- here("Figures")
```


```{r}
bnf <- read.csv(here(data_loc,'BNF_Sampled.csv'), header= TRUE)
denit <- read.csv(here(data_loc,'Denitrification_Sampled.csv'), header= TRUE)
```


##IGBP Bar Chart 

```{r}
igbp_rough_sample <- read.csv(here(data_loc, 'IGBP_Plot.csv'), header = TRUE)
```



```{r}
colors <- c("grey", "red", "blue")  # Replace with your desired colors

# Define the order of categorical variables
custom_order <- c("ENF", "EBF", "DNF", "DBF", "Mixed Forests", "Savannas", "Grassland", "Permanent Wetlands", "Barren")  # Replace with your desired order
igbp_rough_sample$SampleRate_BNF_sites
# Create a grouped bar plot using specific columns as y-values
p <- plot_ly(igbp_rough_sample, x = ~Biome_rough, type = NULL, marker = list(opacity = 0.7)) %>%
  add_trace(y = ~SampleRate_BNF_measurements, name = "BNF", type = "bar", marker = list(color = colors[2])) %>%
  add_trace(y = ~SampleRate_denit_measurements, name = "Denitrification", type = "bar", marker = list(color = colors[3])) %>%
  layout(title = "IGBP Coarse-Scale Land Cover Difference",
         xaxis = list(title = "Biome", categoryorder = "array", categoryarray = custom_order),
         yaxis = list(title = "Percent difference (%)"),
         barmode = "group",
         legend = list(x = 0.9, y = 0.9))

# Print the updated plot
p
```

```{r}
label_mapping <- c("EBF" = "Evergreen Broadleaf Forest",
                   "DBF" = "Deciduous Broadleaf Forest",
                   "Mixed Forests" = "Mixed Forests",
                   "Savannas" = "Savannas",
                   "Grassland" = "Grassland",
                   "Permanent Wetlands" = "Permanent Wetlands",
                   "Barren" = "Barren")
igbp_rough_sample$CustomLabels <- label_mapping[igbp_rough_sample$Biome_rough]

p <- plot_ly(igbp_rough_sample, x = ~CustomLabels, type = NULL, marker = list(opacity = 0.7)) %>%
  add_trace(y = ~SampleRate_BNF_sites, name = "BNF", type = "bar", marker = list(color = colors[2])) %>%
  add_trace(y = ~SampleRate_denit_sites, name = "Denitrification", type = "bar", marker = list(color = colors[3])) %>%
  layout(title = " ",
         xaxis = list(title = " ", categoryorder = "array", categoryarray = unique(igbp_rough_sample$CustomLabels), tickangle = 30),
         yaxis = list(title = " "),
         barmode = "group",
         legend = list(x = 0.9, y = 0.9))
p


# Create annotations for percentage labels
annotations <- list()
for (i in 1:nrow(igbp_rough_sample)) {
  # Find the max absolute value for correct positioning
  max_val <- max(abs(igbp_rough_sample$SampleRate_BNF_sites[i]), abs(igbp_rough_sample$SampleRate_denit_sites[i]))
  
  # Dynamic offset (5% of max value)
  y_offset <- max_val * 0.05  

  # Determine which bar is taller (absolute value)
  if (abs(igbp_rough_sample$SampleRate_BNF_sites[i]) >= abs(igbp_rough_sample$SampleRate_denit_sites[i])) {
    y_pos <- igbp_rough_sample$SampleRate_BNF_sites[i] + ifelse(igbp_rough_sample$SampleRate_BNF_sites[i] > 0, y_offset, -y_offset)
    y_anchor <- ifelse(igbp_rough_sample$SampleRate_BNF_sites[i] > 0, 'bottom', 'top')
  } else {
    y_pos <- igbp_rough_sample$SampleRate_denit_sites[i] + ifelse(igbp_rough_sample$SampleRate_denit_sites[i] > 0, y_offset, -y_offset)
    y_anchor <- ifelse(igbp_rough_sample$SampleRate_denit_sites[i] > 0, 'bottom', 'top')
  }
  
  # Add annotation for each bar
  annotations <- c(annotations, list(
    list(
      x = igbp_rough_sample$CustomLabels[i],  # Correct x-axis mapping
      y = y_pos,
      text = paste0(igbp_rough_sample$Percent_rough_total[i], "%"),
      showarrow = FALSE,
      xanchor = 'center',
      yanchor = y_anchor
    )
  ))
}

# Add annotations to the plot
p <- layout(p, annotations = annotations)

# Print the updated plot
p


```


##Density Plots

Soil Phosphorous 
```{r}
SoilP <- raster(here(rast_loc,'soil_total_phosphorus_Clipped.tif'))

# plot raster data
plot(SoilP,
     main = "Soil P")
```

Convert raster values to data frame
```{r}
SoilP <- brick(here(rast_loc,'soil_total_phosphorus_Clipped.tif'))
```

```{r}
SoilP_df <- as.data.frame(rasterToPoints(SoilP,spatial = TRUE))
head(SoilP_df)
hist(SoilP_df$soil_total_phosphorus_Clipped)
```
```{r}
hist(bnf$soil_total_phosphorus_EX)
```
```{r}
hist(denit$Extracted_soil_total_phosphorus)
```


```{r}
# Set your desired x-axis extent
x_axis_extent <- c(0,1500)

# Specify the colors for each dataset
color_palette <- c("Soil P - Tropics" = "grey", "Soil P - Fixation points" = "red", "Soil P - Denitrification points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
p <-ggplot() +
  geom_density(data = SoilP_df, aes(x = soil_total_phosphorus_Clipped, fill = "Soil P - Tropics"), alpha = 0.5) +
  geom_density(data = bnf, aes(x = soil_total_phosphorus_EX, fill = "Soil P - Fixation points"), alpha = 0.5) +
  geom_density(data = denit, aes(x = Extracted_soil_total_phosphorus, fill = "Soil P - Denitrification points"), alpha = 0.5) +
  labs(title = " ", x = "Soil P (mg/kg)", y = " ") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(
    legend.position = c(0.83, 0.86), # Adjust legend position
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14)   # Adjust font size for axis text
  )
p
ggsave(here(figure_loc, "SoilP_Density.png"), plot = p, width = 8, height = 6, dpi = 300)
```

Density curves with lines indicating the peak (mode) of each population
```{r}
# --- 1. Calculate modes (peak positions) for each group ---
# Remove NAs if needed
mode_tropics <- density(na.omit(SoilP_df$soil_total_phosphorus_Clipped))$x[which.max(density(na.omit(SoilP_df$soil_total_phosphorus_Clipped))$y)]
mode_bnf     <- density(na.omit(bnf$soil_total_phosphorus_EX))$x[which.max(density(na.omit(bnf$soil_total_phosphorus_EX))$y)]
mode_denit   <- density(na.omit(denit$Extracted_soil_total_phosphorus))$x[which.max(density(na.omit(denit$Extracted_soil_total_phosphorus))$y)]

# --- 2. Set axis limits and color palette ---
x_axis_extent <- c(0, 1400)

color_palette <- c(
  "Soil P - Tropics" = "grey",
  "Soil P - Fixation points" = "red",
  "Soil P - Denitrification points" = "blue"
)

# --- 3. Create the plot with density curves and vertical dashed lines ---
p <- ggplot() +
  geom_density(data = SoilP_df, aes(x = soil_total_phosphorus_Clipped, fill = "Soil P - Tropics"), alpha = 0.5) +
  geom_density(data = bnf, aes(x = soil_total_phosphorus_EX, fill = "Soil P - Fixation points"), alpha = 0.5) +
  geom_density(data = denit, aes(x = Extracted_soil_total_phosphorus, fill = "Soil P - Denitrification points"), alpha = 0.5) +
  
  # Add vertical dashed lines at density peaks (modes)
  geom_vline(xintercept = mode_tropics, linetype = "dashed", color = color_palette["Soil P - Tropics"], linewidth = 0.75) +
  geom_vline(xintercept = mode_bnf,     linetype = "dashed", color = color_palette["Soil P - Fixation points"], linewidth = 0.75) +
  geom_vline(xintercept = mode_denit,   linetype = "dashed", color = color_palette["Soil P - Denitrification points"], linewidth = 0.75) +

  # Labels, theme, and scale adjustments
  labs(title = " ", x = "Soil P (mg/kg)", y = " ") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +
  theme(
    legend.position = c(0.83, 0.86),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )

# Show plot
p
#ggsave(here(figure_loc, "SoilP_Density.png"), plot = p, width = 8, height = 6, dpi = 300)
```





##Density curve offset 
BNF - Differences in curve modes -> peak vlaues and offset 
```{r}
sub_values <- bnf$soil_total_phosphorus_EX
sub_values_clean <- sub_values[!is.na(sub_values) & sub_values > 0]

pop_values <- SoilP$soil_total_phosphorus_Clipped
pop_values_clean <- pop_values[!is.na(pop_values) & pop_values > 0]

# Estimate densities
pop_density <- density(pop_values_clean, from = 0, to = 1400, n = 512)
sub_density <- density(sub_values_clean, from = 0, to = 1400, n = 512)

# Find peak x-value (mode) for each curve
pop_peak_x <- pop_density$x[which.max(pop_density$y)]
sub_peak_x <- sub_density$x[which.max(sub_density$y)]

# Compute offset directly from peaks
peak_offset <- sub_peak_x - pop_peak_x

# Report results
cat("Peak (mode) of population density:", pop_peak_x, "\n")
cat("Peak (mode) of subset density:", sub_peak_x, "\n")
cat("Offset between peaks:", peak_offset, "units of phosphorus\n")
```


Cross correlation method, non-parametric, robust 
```{r}
# Estimate KDEs for both datasets over the same fixed grid
x_grid <- seq(0, 1400, length.out = 512)



# Kernel density estimates
pop_density <- density(SoilP$soil_total_phosphorus_Clipped, from = 0, to = 1400, n = length(x_grid))
sub_density <- density(sub_values_clean, from = 0, to = 1400, n = length(x_grid))

# Perform cross-correlation on the y-values of the densities
ccf_result <- ccf(pop_density$y, sub_density$y, plot = TRUE)

# Calculate the offset in original units (phosphorus units)
dx <- diff(x_grid[1:2])
offset_index <- ccf_result$lag[which.max(ccf_result$acf)]
offset_value <- offset_index * dx

cat("Estimated offset (cross-correlation):", offset_value, "units of phosphorus\n")
```
Negative offset means the second whole curve is offset to the left, positive is to the right 

```{r}
# Spacing between x-values in your density grid
dx <- diff(x_grid[1:2])  # This gives the size of one step in phosphorus units

# Index of max correlation
best_lag_index <- which.max(ccf_result$acf)
best_lag <- ccf_result$lag[best_lag_index]

# Convert lag index to actual phosphorus offset
offset_value <- best_lag * dx

cat("Estimated shift (subset vs population):", offset_value, "mg/kg P")
```





Denit: Differences in curve modes -> peak vlaues and offset 
```{r}
sub_values <- denit$Extracted_soil_total_phosphorus
sub_values_clean <- sub_values[!is.na(sub_values) & sub_values > 0]

pop_values <- SoilP$soil_total_phosphorus_Clipped
pop_values_clean <- pop_values[!is.na(pop_values) & pop_values > 0]

# Estimate densities
pop_density <- density(pop_values_clean, from = 0, to = 1400, n = 512)
sub_density <- density(sub_values_clean, from = 0, to = 1400, n = 512)

# Find peak x-value (mode) for each curve
pop_peak_x <- pop_density$x[which.max(pop_density$y)]
sub_peak_x <- sub_density$x[which.max(sub_density$y)]

# Compute offset directly from peaks
peak_offset <- sub_peak_x - pop_peak_x

# Report results
cat("Peak (mode) of population density:", pop_peak_x, "\n")
cat("Peak (mode) of subset density:", sub_peak_x, "\n")
cat("Offset between peaks:", peak_offset, "units of phosphorus\n")
```

Cross correlation method, non-parametric, robust 
```{r}
# Estimate KDEs for both datasets over the same fixed grid
x_grid <- seq(0, 1400, length.out = 512)

# Kernel density estimates
pop_density <- density(SoilP$soil_total_phosphorus_Clipped, from = 0, to = 1400, n = length(x_grid))
sub_density <- density(sub_values_clean, from = 0, to = 1400, n = length(x_grid))

# Perform cross-correlation on the y-values of the densities
ccf_result <- ccf(pop_density$y, sub_density$y, plot = TRUE)

# Calculate the offset in original units (phosphorus units)
dx <- diff(x_grid[1:2])
offset_index <- ccf_result$lag[which.max(ccf_result$acf)]
offset_value <- offset_index * dx

cat("Estimated offset (cross-correlation):", offset_value, "units of phosphorus\n")
```

```{r}
# Spacing between x-values in your density grid
dx <- diff(x_grid[1:2])  # This gives the size of one step in phosphorus units

# Index of max correlation
best_lag_index <- which.max(ccf_result$acf)
best_lag <- ccf_result$lag[best_lag_index]

# Convert lag index to actual phosphorus offset
offset_value <- best_lag * dx

cat("Estimated shift (subset vs population):", offset_value, "mg/kg P")
```




##Checking colinearity of environemntal factors
```{r}
#remove NA values and test for colinearity 
 raster_data[raster_data==0] <- NA
 raster_data
 
 raster_data_noZeros<-raster_data[complete.cases(raster_data),]


# We can visually look for correlations between variables:
heatmap(abs(cor(raster_data_noZeros)), 
        # Compute pearson correlation (note they are absolute values)
        col = rev(heat.colors(6)), 
        Colv = NA, Rowv = NA)
legend("topright", 
       title = "Absolute Pearson R",
       legend =  round(seq(0,1, length.out = 6),1),
       y.intersp = 0.7, bty = "n",
       fill = rev(heat.colors(6)))
```



##Principal Component Analysis 
only using those factors that have an absolute pearson R less than 0.8 
```{r}
data <- read.csv(here(data_loc,'RandPoints_Ext.csv'), header= TRUE)
data$Slope <- as.numeric(data$Slope)
data <- data %>% dplyr::select(-c(NEE, IGBP, GPP, Elevation, Roughness, SoilpH, PPT))

#data <- sapply( data, as.numeric, rm.na = TRUE)

```

```{r}
for (col in 1:ncol(data)) {
    hist(data[,col], main=colnames(data)[col],  xlab=colnames(data)[col])
}
```


Normalizing values 
```{r}


#data$PPT <- (data$PPT) +0.001

#data$PPT <- log(data$PPT)

data$MAT <- (data$MAT) +0.001

data$MAT <- log(data$MAT)

#data$GPP <- (data$GPP) +0.001

#data$GPP <- log(data$GPP)

data$NPP <- (data$NPP) +0.001

data$NPP <- log(data$NPP)

data$Ndep2050 <- (data$Ndep2050) +0.001

data$Ndep2050 <- log(data$Ndep2050)

data$Slope <- (data$Slope) +0.001

data$Slope <- log(data$Slope)

#data$Elevation <- (data$Elevation) +0.001

#data$Elevation <- log(data$Elevation)

#data$Roughness <- (data$Roughness) +0.001

#data$Roughness <- log(data$Roughness)

data$SoilOC <- (data$SoilOC) +0.001

data$SoilOC <- log(data$SoilOC)

data$SoilN <- (data$SoilN) +0.001

data$SoilN <- log(data$SoilN)

data$SoilP <- (data$SoilP) +0.001

data$SoilP <- log(data$SoilP)

#data$SoilpH <- (data$SoilpH) +0.001

#data$SoilpH <- log(data$SoilpH)

data$DtB1 <- (data$DtB1) +0.1

data$DtB1 <- log(data$DtB1)
```

```{r}
for (col in 1:ncol(data)) {
    hist(data[,col], main=colnames(data)[col],  xlab=colnames(data)[col])
}
```

Based on QCBS vegan notes (https://r.qcbs.ca/workshop09/book-en/learning-objectives.html)
FactoMineR documentation

```{r}
# Frist, normalize and then standardize your data (data previously normalized and labels removed)

#soil_var.2 <- decostand(soil_var.2, "standardize")

raster_data <- decostand(data, "standardize")
```

```{r}
# Perform PCA with FactoMineR 
# visualize with factoextra

raster.pca <- FactoMineR::PCA(raster_data)
 
summary(raster.pca)  

  ## unknown scaling for the below plot
```

```{r}
# Control variable colors using their contribution
# Possible values for the argument col.var are :
  # "cos2", "contrib", "coord", "x", "y"
fviz_pca_var(raster.pca, col.var="contrib") +
scale_color_gradient2(low="blue", mid="yellow", 
                      high="orange", midpoint=5)+theme_bw()
```

```{r}
raster.pca.plot <- factoextra::fviz_pca_biplot(raster.pca, # PCA object
                axes = c(1, 2), # graph first two PCs
                geom.ind = "point", # display trees as points
                geom.var = c("arrow", "text"), # add arrows and text to variables
                pointsize = 2, # increase size of tree points
                labelsize = 6, # increase graph label font size
                alpha = 0.5, # fade points/ellipses to read better
                arrowsize = 1,
                #habillage = as.factor(soil_samp$site), 
                # habillage = as.factor(soil_clusters),
                addEllipses = FALSE, # draw ellipses around groups (sites)
                ellipse.level = 0.95,
                #palette = pal3, # match palette with other plots
                col.ind ="grey",
                col.var = "blue", # make arrows/text of var black
                ggrepel = TRUE, # avoid label overlap
                mean.point = FALSE) + # remove site average point
        theme(panel.grid = element_blank(),
        legend.title = element_blank()) +
  geom_text(aes(x= -5, y = 4), label = "", size = 4) +
  theme(text = element_text(size = 17),
        axis.title = element_text(size = 17),
        axis.text = element_text(size = 17),
        title = element_blank()) 

raster.pca.plot

# export plot to output folder

#png("All_PCA_2024-4-09.png", height = 8, width = 12, unit = "in", res = 800)
#print(raster.pca.plot)
#dev.off()
```

```{r}
# obtain correlation coefficients between variables and
# PCs, and run a significance test on correlations

FactoMineR::dimdesc(raster.pca, axes = c(1, 2))
    ## all variables are significantly correlated with PCS
```

```{r}
## examine variables loadings for PCs
## A reference dashed line is also shown on the barplot. 
## This reference line corresponds to the expected value 
## if the contribution where uniform. Bars above this dashed line are important contributors 

fviz_contrib(raster.pca, choice = "var", axes = 1)
fviz_contrib(raster.pca, choice = "var", axes = 2)
fviz_contrib(raster.pca, choice = "var", axes = 3)  
fviz_contrib(raster.pca, choice = "var", axes = 4) 

# rm(soil.pca.plot, soil.pca)

```

```{r}

# Kaiser Guttman criteria - select PCs that explain the more than
# the average across all PCs

ev <- raster.pca$eig[ ,1] # select eigenvalues for each variable
barplot(ev)
abline(h = mean(ev), col = "red3")
legend("topright", "Average Eigenvalue", col = "red3")

# useful information down to PC3

```

```{r}
#Make a scree plot using base graphics : A scree plot is a graph of the eigenvalues/variances associated with components.
eigenvalues <- raster.pca$eig
barplot(eigenvalues[, 2], names.arg=1:nrow(eigenvalues), 
       main = "Variances",
       xlab = "Principal Components",
       ylab = "Percentage of variances",
       col ="steelblue")
# Add connected line segments to the plot
lines(x = 1:nrow(eigenvalues), eigenvalues[, 2], 
      type="b", pch=19, col = "red")
```

##Reattaching locations and categorical variables to PC scores 
```{r}
pc_scores <- as.data.frame(raster.pca$ind$coord)
```

```{r}
data2 <- read.csv(here(pca_loc, 'RandPoints_Ext_locations.csv'), header= TRUE)
```

```{r}
with_loc <- merge(data2, pc_scores, by = 0)
#write.csv(with_loc, file = "PCScores_noColinear_5-10-24.csv")
```

```{r}
with_loc <- subset(with_loc, Dim.2 < 20, select = Row.names:Dim.5)
```

```{r}
cols <- c("#F72424", "#0000FF", "#A0A0A050")
ordered <- with_loc[order(with_loc$Type, decreasing = TRUE), ]

points <- ggplot(ordered, aes(x = Dim.1, y = Dim.2, color = Type, order = Type)) +
  geom_hline(yintercept = 0, linetype = "solid", size = 1, colour = "black") +  # Bold zero horizontal line
  geom_vline(xintercept = 0, linetype = "solid", size = 1, colour = "black") +
  geom_point() +
  scale_color_manual(values = cols) + 
  theme_bw()+
  xlab("PC1") + ylab("PC2")
   # Bold zero vertical line
points

#png("PC_Scores_PointsOver_Cropped.png", height = 8, width = 12, unit = "in", res = 800)
#print(points)
#dev.off()
```

##Generalized Additive Model (GAM)
```{r}
df <- read.csv(here(pca_loc,'RandPoints_Ext.csv'), header= TRUE)
df$Slope <- as.numeric(df$Slope)
df <- df %>% dplyr::select(-c(NEE, IGBP, GPP, Elevation, Roughness, SoilpH, PPT))
```



```{r}
#read in data and standardize variables 
numeric_columns <- sapply(df, is.numeric)
df[numeric_columns] <- scale(df[numeric_columns])
```

```{r}
#attach them to the presence/absence classification
gam_data <- read.csv(here(pca_loc,"RandPoints_Ext_all_GAM_loc.csv"), header = TRUE)
gam_data <- merge(gam_data, df, by = 0)
gam_data <- gam_data %>% dplyr::select(-c(Row.names))
```

```{r}
#remove NA rows 
gam_data <- na.omit(gam_data)
```

```{r}
hist(gam_data$PresAbs)
```

```{r}
#try to downsample data 
##Remove pres abs data rows first 
# Specify the column and value to remove
column_name <- "PresAbs"
value_to_remove <- "1"

# Separate the rows to be removed
rows_to_remove <- gam_data[gam_data[[column_name]] == value_to_remove, ]

# Data frame without the specified rows
df_remaining <- gam_data[gam_data[[column_name]] != value_to_remove, ]
```

```{r}
#actual downsample
if (nrow(df_remaining) >= 10000) {
  # Randomly sample 10,000 rows
  sampled_df <- df_remaining[sample(nrow(df_remaining), 10000), ]
} else {
  stop("Not enough rows with the specified value to sample 10,000 rows.")
}
```

```{r}
#reattach 
gam_data <- rbind(sampled_df, rows_to_remove)
```

Doing the GAM with 10,000 samples and considering potentially interesting variables: 
```{r}
#try the Gam

gam_model<-gam(PresAbs~ SoilN+SoilOC+NPP+MAT+Slope+SoilP+AET+SoilMoisture+Seasonality,family = binomial, data=gam_data, select = TRUE, method="GCV.Cp")
#Then when you want to predict, you just:
Gam_pred <- predict.gam(tmp_sdm, newdata=gam_data, type="response")
Gam_pred <- as.data.frame(Gam_pred)
#predictions for presence absence at each point based only on PPT and MAT

```

```{r}
hist(Gam_pred$Gam_pred)
```

```{r}
summary(gam_model)
#already cross validated 
```

```{r}
gam.check(gam_model)
```


```{r}
#Cross-validation
# Perform k-fold cross-validation
gam_data$PresAbs <- as.factor(gam_data$PresAbs)
train_control <- trainControl(method = "cv",
                              number = 10)
cv_model <- train(PresAbs ~ SoilN+SoilOC+NPP+MAT+Slope+SoilP+DtB1+AET+SoilMoisture+Seasonality, data = gam_data, method = "gam", trControl = train_control)
print(cv_model)
```
Accuracy indicates that the model did well to predict negative and positive values. Kapps indicates there is fair agreement, I think meaning that our model is 25% better than what would align only by chance. 


Attaching the GAM results to locaitons
```{r}
gam_data <- merge(gam_data, Gam_pred, by = 0)
gam_data <- gam_data %>% dplyr::select(-c(Row.names))
```

```{r}
file_path <- here(data_loc, "Gam_10000_output.csv")
write.csv(gam_data, file = file_path, row.names = FALSE)
```



##TRENDY-N Models 
```{r}
loc <- here("Data", "TRENDYN")
CP_NMin <- rast(here(loc, "CABLE-POP/fNnetmin_annually.nc"))
CLASSIC_NMin <- rast(here(loc,"CLASSIC/fNnetmin_annually.nc"))
CLM5_NMin <- rast(here(loc,"CLM5.0/fNnetmin_annually.nc"))
DLEM_NMin <- rast(here(loc,"DLEM/fNnetmin_annually.nc"))
#ISAM_NMin <- rast(here(loc,"ISAM/fNnetmin_annually.nc"))
JSBACH_NMin <- rast(here(loc,"JSBACH/fNnetmin_annually.nc"))
JULES_ES_NMin <- rast(here(loc,"JULES-ES/fNnetmin_annually.nc"))
LPJ_GUESS_NMin <- rast(here(loc,"LPJ-GUESS/fNnetmin_annually.nc"))
LPX_Bern_NMin <- rast(here(loc,"LPX-Bern/fNnetmin_annually.nc"))
OCNv2_NMin <- rast(here(loc,"OCNv2/fNnetmin_annually.nc"))
ORCHIDEEv3_NMin <- rast(here(loc,"ORCHIDEEv3/fNnetmin_annually.nc"))
```

BNF - Run this block of code for each individual model, ensuring to change the name of the output file! 
```{r}
Classic_bnf_nc <- ncdf4::nc_open(here(loc, "CLASSIC/fBNF_annually.nc"))

data_var <- ncvar_get(Classic_bnf_nc, "fBNF")
time_var <- ncvar_get(Classic_bnf_nc, "time")

lat <- ncvar_get(Classic_bnf_nc, "latitude")
lon <- ncvar_get(Classic_bnf_nc, "longitude")

nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)

selected_time_steps <- c(150:170)

rasters <- list()

for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}

raster_brick <- brick(rasters)

average_raster <- calc(raster_brick, mean)


##CONVERTS UNITS

average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

values(average_raster_units)

# Define the output file path
output_file <- here(averages_loc,"Classic_bnf_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)

```

```{r}
# Calculate the mean across the raster stack, removing NA values
mean_raster <- app(spat_raster_stack, fun = function(x) mean(x, na.rm = TRUE))
output_file <- here("Outputs", "Mean_TRENDYN_BNF.tif")
writeRaster(mean_raster, output_file, overwrite = TRUE)

# Calculate the standard deviation across the raster stack, removing NA values
sd_raster <- app(spat_raster_stack, fun = function(x) sd(x, na.rm = TRUE))
output_file <- here("Outputs","SD_TRENDYN_BNF.tif")
writeRaster(sd_raster, output_file, overwrite = TRUE)

# Calculate the coefficient of variation (CV)
cv_raster <- sd_raster / mean_raster
output_file <- here("Outputs","CV_TRENDYN_BNF.tif")
writeRaster(cv_raster, output_file, overwrite = TRUE)

# Plot the CV raster
plot(sd_raster)
```


N loss - Run this block of code for each individual model, ensuring to change the name of the output file!
```{r}
#averages_loc_nloss <- here("Outputs", "AvgRasters", "Nloss")

Classic_nloss_nc <- ncdf4::nc_open(here(loc, "ORCHIDEEv3/fNloss_annually.nc"))

data_var <- ncvar_get(Classic_nloss_nc, "fNloss")
time_var <- ncvar_get(Classic_nloss_nc, "time")

lat <- ncvar_get(Classic_nloss_nc, "latitude")
lon <- ncvar_get(Classic_nloss_nc, "longitude")

nlat <- length(lat)
nlon <- length(lon)
ntime <- length(time_var)

selected_time_steps <- c(150:170)

rasters <- list()

for (i in selected_time_steps) {
  data_slice <- data_var[,,i]
  r <- raster(t(data_slice), xmn=min(lon), xmx=max(lon), ymn=min(lat), ymx=max(lat), crs=CRS("+proj=longlat +datum=WGS84"))
  rasters[[length(rasters) + 1]] <- r
}

raster_brick <- brick(rasters)

average_raster <- calc(raster_brick, mean)




average_raster <- flip(average_raster, direction = 'y')
average_raster <- terra::rotate(average_raster)
average_raster_units <- average_raster*(3.1104e+10)
plot(average_raster_units)
#average_raster

values(average_raster_units)

# Define the output file path
output_file <- here(averages_loc_nloss,"ORCHIDEEv3_Nloss_Avg.tif")

# Write the adjusted average raster to a GeoTIFF file
writeRaster(average_raster_units, output_file, format = "GTiff", overwrite = TRUE)

```

```{r}
# Calculate the mean across the raster stack, removing NA values
mean_raster <- app(spat_raster_stack, fun = function(x) mean(x, na.rm = TRUE))
output_file <- "Mean_TRENDYN_Nloss.tif"
writeRaster(mean_raster, output_file, overwrite = TRUE)

# Calculate the standard deviation across the raster stack, removing NA values
sd_raster <- app(spat_raster_stack, fun = function(x) sd(x, na.rm = TRUE))
output_file <- "SD_TRENDYN_Nloss.tif"
writeRaster(sd_raster, output_file, overwrite = TRUE)

# Calculate the coefficient of variation (CV)
cv_raster <- sd_raster / mean_raster
output_file <- "CV_TRENDYN_Nloss.tif"
writeRaster(cv_raster, output_file, overwrite = TRUE)

# Plot the sd raster
plot(sd_raster)
```


##Plotting density curves with TRENDY data
```{r}
bnf <- read.csv(here(data_loc,'BNF_TRENDYN.csv'), header= TRUE)
bnf_no_NA <- bnf %>% drop_na(Variance)

denit <- read.csv(here(data_loc,'Nloss_TRENDYN.csv'), header= TRUE)
denit_no_NA <- denit %>% drop_na(Variance)
```

BNF Density curves - CV
```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_bnf <- brick(here(outputs_loc,"Tropics_BNF_CV.tif"))
cv_bnf_df <- as.data.frame(rasterToPoints(cv_bnf,spatial = TRUE))
```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_bnf_global <- brick(here(averages_loc, "BNF_CV.tif"))
cv_bnf_global_df <- as.data.frame(rasterToPoints(cv_bnf_global,spatial = TRUE))
```


```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("BNF - Tropics" = "grey", "BNF - Sampled points" = "red")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
p <-ggplot() +
  geom_density(data = cv_bnf_df, aes(x = lyr.1, fill = "BNF - Tropics"), alpha = 0.5) +
  geom_density(data = bnf_no_NA, aes(x = CV, fill = "BNF - Sampled points"), alpha = 0.5) +
  labs(title = " ", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
  theme(
    legend.position = c(0.83, 0.86), # Adjust legend position
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14)   # Adjust font size for axis text
  )
p
#ggsave("CV_BNF_Clean.png", plot = p, width = 8, height = 6, dpi = 300)
```

Denit Density curves - CV
```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_denit <- brick(here(outputs_loc,"Tropics_Nloss_CV.tif"))
cv_denit_df <- as.data.frame(rasterToPoints(cv_denit,spatial = TRUE))
```

```{r}
#model_bnf <- raster(x = "CV_TRENDYN_BNF.tif")
cv_denit_global <- brick(here(averages_loc,"Nloss_CV.tif"))
cv_denit_global_df <- as.data.frame(rasterToPoints(cv_denit_global,spatial = TRUE))
```

```{r}
# Assuming df1, df2, and df3 are your dataframes with the respective columns

# Install and load necessary libraries if not already installed
# install.packages("ggplot2")
# library(ggplot2)

# Set your desired x-axis extent
x_axis_extent <- c(0,3)

# Specify the colors for each dataset
color_palette <- c("N loss - Tropics" = "grey", "Denit - Sampled points" = "blue")

# Create density plots using ggplot2 with specified x-axis extent, colors, and legend
p <- ggplot() +
  geom_density(data = cv_denit_global_df, aes(x = lyr.1, fill = "N loss - Tropics"), alpha = 0.5) +
  geom_density(data = denit_no_NA, aes(x = CV, fill = "Denit - Sampled points"), alpha = 0.5) +
  labs(title = " ", x = "Coefficient of Variation", y = "Density") +
  theme_minimal() +
  xlim(x_axis_extent) +
  scale_fill_manual(values = color_palette, name = " ") +  # Set legend name and colors
   theme(
    legend.position = c(0.83, 0.86), # Adjust legend position
    axis.title = element_text(size = 16), # Adjust font size for axis titles
    axis.text = element_text(size = 14)   # Adjust font size for axis text
  )
p
#ggsave("CV_NLoss_Clean.png", plot = p, width = 8, height = 6, dpi = 300)
```

















